Review Paper Draft

### 2020-07-01 main.tex
4.1 - 4.2 edits

We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations. \anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership \pb{Is this true for the linear mixing case as well?} \anna{yes}. 1000 simulations were performed for each data type and parameter combination.

@@

We begin by comparing the ordinary lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied. Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) in various settings and broken down into bias and variance components.  Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$. The top row of figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-Lasso perform comparably. PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors have been added to the model, but none of them actually explain any variability in the outcome. The bottom row of figure \ref{fig:mse} shows the MSE when $\xi = 0.8$ (i.e., 40\% of the variability in the phenotype is attributable to environmental confounding). In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso. This improvement is more substantial in the independent subpopulations simulated data due to the fact that the subpopulations are more easily distinguishable based on their observed genetic data.

\pb{Need to describe how we select $\lambda$}

\anna{
\begin{itemize}
    \item Need details about the real data once it's run
    \item Need to fix facet labels
\end{itemize}
}

@@

\pb{Just FYI, it's hard to tell bias and var apart in black and white}

@@

Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example, and (2) a large number of small, closely related subpopulations to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms.

>>>>

We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations. \anna{cite Vanderwas data} $p = 1000$ for all three data types, $n = 200$ for the simulated data types, and 50 causal SNPs were used throughout. The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method. For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership, \anna{or founding population membership in the case of the admixed data.} 1000 simulations were performed for each data type and parameter combination.

@@

We begin by comparing the ordinary lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied. Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) in various settings and broken down into bias and variance components.  Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$. The top row of figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-Lasso perform comparably. PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors have been added to the model, but none of them actually explain any variability in the outcome. The bottom row of figure \ref{fig:mse} shows the MSE when $\xi = 0.8$. In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso. This improvement is more substantial in the independent subpopulations simulated data than the admixed simulated data. This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.

@@

\pb{Just FYI, it's hard to tell bias and var apart in black and white} \anna{will fix this and greek letters}

@@

Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) coarse \anna{cut this?: relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example,} and (2) fine \anna{cut?: a large number of small subpopulations corresponding to finer grained environmental heterogeneity, such as household exposures or behavioral norms}.
